---
title: "CHEMTAX Main"
author: "Sebastian DiGeronimo"
date: "6/2/2022"
output:
  html_document: default
  pdf_document: default
---

### Intro: 
This will serve as my starting point for converting CHEMTAX V2.0 into R.
#TODO: fix this: [CHEMTAX Document](./Re__CHEMTAX_Update_[SEC=OFFICIAL]/Re__CHEMTAX_Update_[SEC=OFFICIAL]/CHEMTAX 2 Octave version.docx)

`r gsub("\\.m$","", list.files(path = "./Re__CHEMTAX_Update_[SEC=OFFICIAL]/MatrixFactorsOctave/", pattern="\\.m$"))`
# Files that need to be converted
double tilde ~ will show done

* "adddir"
* "amatfactsd"                    ---- started
* "bmatfactsd"                    ---- started         
* "bootln"                        ---- started * connected to `nnmatfactsd`           
* "bootnp"              
* "brokewest"          
* "chemtaxbrokewest"    
* "CHEMTAXBROKEWestb"   
* "CHEMTAXBROKEWests"   
* "CHEMTAXBROKEWestx"   
* "dataout"             
* "Formulas"           
* "initstruct"          
* "matfactuvw"          
* "nnmatfactsd"         
* "normprod"            
* "permcalc"            
* "randstart"          
* "regplot"             
* "saz"                 
* "sazchemtax090210"    
* "SAZS_CHEMTAX090210s"


#TODO: something to include is:
This would load internal ratios file so that you have a basic pigment ratio var
pigment_ratio <- system.file("", package = "rCHEMTAX")

## Section 1: Setup

```{r setup, include=FALSE, message=FALSE}
root <- rprojroot::find_rstudio_root_file()
library("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = root)
library("tidyverse")
library("dplyr")
library("lubridate")
library("vroom")


```

```{r create-dir}
# will create a set directory if does not exists
# useful for new projects
mainDir <- getwd()
subDir <-
    c("data/raw",
      "data/processed",
      "data/plots",
      "data/metadata",
      "Rmd",
      "scripts")

fs::dir_create(path = paste0(mainDir,"/",subDir))
rm(mainDir, subDir)
```


```{r}

```

