---
title: "CHEMTAX Data Test"
author: "Sebastian DiGeronimo"
date: "2023-01-25"
output: html_document
---
# ---- Load functions and libraries ----
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("here")
library("pracma") # functions imported from MatLab
library("rprojroot")
library("ggplot2")
library("tictoc")
library("stringr")
library("limSolve")
library("cli")
library("readxl")
library("tidyr")
library("rlang")
library("ggforce")
library("dplyr")


# source scripts
# source(here("scripts", "chemtaxbrokewest.R")) # load stock data the old way
source(here("scripts", "load_pigment_data.R"))  # load stock data
source(here("scripts", "nnmatfactsd.R"))        # non-neg matrix factorization
source(here("scripts", "normprod.R"))           # normalize outputs
source(here("scripts", "bmatfactsd.R"))         # calc F matrix from a_est
source(here("scripts", "regplot.R"))            # reglarization test
source(here("scripts", "randstart.R"))          # randomization test
source(here("scripts", "bootln.R"))             # log
source(here("scripts", "bootnp.R"))             # bootstrap
```
# ---- Load data ----
Select between two datasets. Either Brokewest or Saz.
```{r load}
# select dataset: brokewest or saz datasets
data_set <- "saz" # 'broke' or 'saz'

if (data_set == "broke") {
  # Brokewest
  pig_dat_file <- here("data", "raw", "CHEMTAXBROKEWests.csv")
  p_ratio_file <- here("scripts", "brokewest_pigment_ratios.csv")
  
  # index of pigments to use, needs to be the same length as F matrix
  idx          <- c(1, 0, 0, 1, 1, 0, 1, 1, 1, 1, 1, 0, 1, 1, 0, 0, 1, 0, 1)
  
} else if (data_set == "saz") {
  # SAZ
  pig_dat_file <- here("data", "raw", "SAZS_CHEMTAX090210s.csv")
  p_ratio_file <- here("scripts", "saz_pigment_ratios.csv")
  
  # index of pigments to use, needs to be the same length as F matrix
  idx          <- c(1, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1)
  
}

# load data set
load_dat <- load_data(
  .file     = pig_dat_file,
  .pig_file = p_ratio_file,
  idx,
  type      = "sd" # change to either sd or norm
)

# parse each matrix to a variable
s     <- load_dat$df         # pigment data
ssd   <- load_dat$df_sd      # s * 0.01 + 0.0003
f0    <- load_dat$pig_r_init # pigment ratio matrix
fsd   <- load_dat$pig_r_sd   # F0 * 0.1 w/ last col = 0.005
taxa  <- load_dat$taxa       # name of taxa groups
pigm  <- load_dat$pigm_sel   # name of selected pigments
```

# ---- Fit non-negative matrix factorization ----
Note: No randomization test is performed here
```{r nnmf}
# fit the matrix factors
temp <- nnmatfactsd(.df         = s,
                    .df_sd      = ssd,
                    .pig_r_init = f0,
                    .pig_r_sd   = fsd,
                    captures    = FALSE)

# results
taxa_amt <- temp$a
f        <- temp$b
info     <- temp$info
```
# ---- Calculate Pigment Ratio Matrix ----
Uses df, df_sd, taxa contribution matrix, pigment ratio matrix, and it sd 
```{r calc-f-mat}
pig_from_df_taxa <- bmatfactsd(
  .df       = s,
  .df_sd    = ssd,
  .taxa_amt = taxa_amt, # output from nnmatfactsd
  .pig_r    = f,
  .pig_r_sd = fsd
)

pig_from_df_taxa
```
# ---- Scale the Factors and Original Data ----
```{r norm}
dat_norm <- normprod(s,taxa_amt,f)
ss       <- temp3$ss
cc       <- temp3$cc
ff       <- temp3$ff
rms      <- temp3$rms
```
# ---- Write Results to File ----
Either:
- broke_processed.csv
- saz_processed.csv
```{r save-result}

if (TRUE) {
  
  file_name <- here("data", "processed", paste0(data_set, "_processed.csv"))
  header    <- paste0(data_set, "\n\n")
  
  # write results to file brokewest.csv
  # pigment ratio
  df1           <-  ff
  colnames(df1) <- pigm
  rownames(df1) <- taxa
  
  # taxa concentration
  df2           <- cc
  colnames(df2) <- taxa

  # Start a sink file with a CSV extension
  sink(file_name)
  
  # Write the first dataframe, with a title and final line separator
  cat(header)
  write.csv(df1, row.names = FALSE)
  cat('\n')
  
  # Write the 2nd dataframe to the same sink
  write.csv(df2, row.names = FALSE)
  
  # Close the sink
  sink()

}
```
# ---- Run Various Test ----
regplot   - test influence of percent error for Std Dev of F matrix
randstart - randomizes the input for either the F matrix, B matrix or F and B
bootln    - convert to log and run test
bootnp    - non-parametric version that subsamples the data each iteration

```{r misc-test}
if (FALSE) {
  # number of replicates
  nrep <- 2
  
  # plot showing the effect of regularization
  # has warning, but seems work
  reg_out <- regplot(s,ssd,f,f, verbose = TRUE)
  
  # show converges from random starts for c
  # sort of works, graphs are not on log scale
  
  rand_out <- 
    randstart(
      s,ssd,f0,fsd,
      .nrep = nrep, .pigm = pigm, .taxa = taxa, 
      verbose = TRUE, .info = list(printitr = 5000)
              )
   
  # bootstrap using parametric log normal
  ln_out <- bootln(
    s,ssd,f0,fsd, 
    .nrep = nrep, .pigm = pigm, .taxa = taxa, 
    verbose = TRUE, .info = list(printitr = 5000)
    )

  # bootstrap non parametric on s
  np_out <- bootnp(
    s,ssd,f0,fsd, 
    .nrep = nrep, .pigm = pigm, .taxa = taxa, 
    verbose = TRUE, .info = list(printitr = 5000)
    )
}
```

